Comprehensive Quality-Aware Automated Semantic Web Service Composition

### 2017-01-29 paper.tex
update

We defined our semantic web service composition solution as a weighted graph $WG = (V, E)$, where $V$ is a set of services as vertex: $V=[S1, S2...Sn]$ and $E$ is a set of edges $E = {e_{1}, e_{2},... e_{n}}$. Each $e$ is associated with $q(e_{sm})$ as weight value that mapped to a pair of quality values $q(e_{mt})$ and $q(e_{s})$ on the edge level, and $e_{m}$ is expressed as $(S_{a},S_{b})={q_{mt}, q_{s}}$. Here we provide an example of the web service composition, which is described in Fig. \ref{wscs}. The data of composition web service flows from the start to the end, where five web services involved and linked with each other using edge connections. Besides that, $q(s_{mt})$ and $q(s_{s})$ are calculated for all edges ${e_{1}, e_{2},... e_{n}}$ assigned as weights.

\begin{figure}[h]
\centering
\fbox{\includegraphics[scale=.29]{weigtedGraphExample.pdf}}
 \caption{Composition weighted graph solution}
 \label{wscs}
\end{figure}

@@

PSO has shown its efficiency in solving combinatorial optimisation problems \cite{fukuyama2008fundamentals}. Therefore, we will employ a PSO-based approach, which is considered to be simple and efficient without penalising or repairing that often required by GP \cite{da2016particle}. Fig. \ref{overview} shows the overview of our approach with five steps. Step 1: The composition process is triggered by a composition goal defined in Subsection \ref{problemDes}, which describes customers requirements both functional and non-functional. Step 2: This composition goal are used to discover all relevant web services, which lead to a shrunken service repository that is subsequently used by PSO as a searching space. Step 3: A weighted graph representation is randomly built up from the an initial service queue that mapped to the particle's location, interleaving with semantic matchmaking process utilising ontology-based index cache. In the weighted graph, graph edges are assigned with semantic matchmaking quality as weights. Step 4: The fitness value of the weighted graph is evaluated to update the position of particle under PSO algorithm in Sect. \ref{pso_algorithm}, where the position is mapped to the index of service queue. later on, the updated service queue is used to decode a new weighted graph as the composition solution. Step 4. Lastly, the best position found in the searching space is selected and decode into the final optimised solution. This PSO-based approach is similar to \cite{da2016particle}, but we employ weighted graphs as a different solution presentation.

\subsection{Ontology-based Index Cached Optimisation}\label{indexCache}
In PSO, particles represent weighted graphs with edges and vertices associated with quality of semantic matchmaking and QoS respectively, the bottlenecks of generating weighted graph lie in building edges and nodes, which are related to the cost of semantic quality calculation and the size of service repository respectively. To effectively construct weighted graphs, we pre-calculate semantic matchmaking quality. The key idea of the index is to create a mapping from the output-related concepts to potentially matched input-related concepts while considering different levels of match types. Meanwhile, this index size could also be reduced by only considering the concepts related in the shrunken service repository, which means the index cache is filtered by those task-relevant web services. This optimised cache also contributes to less and constant time for weighted graphs building through the whole evolutionary process.

@@

\begin{algorithm}
 \setlength\hsize{0.9\linewidth}
 \SetKwInOut{Input}{Input}\SetKwInOut{Output}{Output}
 \SetKwFunction{generateWeightedGraph}{generateWeightedGraph}
 \SetKwProg{Procedure}{Procedure}{}{}
 \LinesNumbered
 \SetNlSty{}{}{:}
 \Input{relevant Web services $rws$}
 \Output{Optimised service queue $queue$,Optimised Weighted Graph $OPTWG$}
 Map each relevant service to an index in the particle's position vector\;
 Randomly initialise each particle in the swarm\;
  \While {max. iterations not met}{
     \ForEach{particles in the swarm}{
     $queue \leftarrow$ particle's position vector\;
     $WG \leftarrow$ $generateWeightedGraph$()\;
	 Calculate the $WG$ fitness value\;

      \eIf{fitness value better than pBest }{    
        Assign current fitness as new pBest\;
       }{
        Keep previous pBest\;
       }	
     }
    Assign best particle's \emph{pBest} value to \emph{gBest}, if better than gBest\;
 	Calculate the velocity of each particle\;
  	Update the position of each particle\;
  }
 \KwRet $OPTWG$\;
 \caption{Steps of the PSO-based Web service composition technique.}
\label{novelSteps}
\end{algorithm}

\begin{algorithm}
 \setlength\hsize{0.9\linewidth}
 \SetKwInOut{Input}{Input}\SetKwInOut{Output}{Output}
 \SetKwFunction{generateWeightedGraph}{generateWeightedGraph}
 \SetKwProg{Procedure}{Procedure}{}{}
 \LinesNumbered
 \SetNlSty{}{}{:}
  \Procedure{\generateWeightedGraph{}}{
 \Input{Task inputs $I$, task outputs $O$, Optimised service queue $queue$, IndexCache $IndexCache$}
 \Output{Weighted Graph $WG$}
 $WG \leftarrow  null$\;
 $WG \leftarrow$ new $endNode$(), new $startNode$()\;
 $OutputSet \leftarrow$ \{$I$\}\;
  \While {all $O$ $\notin$ $OutputSet$  and $queue !=null$}{
     \ForEach{$ws$ in $queue$}{
      \uIf{$ws.inputs$ $\in$ $OutputSet$ }{    
    		 \ForEach{$I$ in $ws.inputs$}{
    		$p_{mt} \leftarrow$ query $IndexCache$\;
     		$p_{s} \leftarrow$ query $IndexCache$\;
              }
        $e_{sm} \leftarrow$ aggregation($p_{mt}$,$p_{s}$)\;
        $WG.edge \leftarrow$  $e_{sm} $\;
        $WG \leftarrow$ new $wsNode$()\;
         $OutputSet$ add \{$ws.outputs$\}\;    
       $queue$.remove $ws$\;
       }	
     }
  }
 remove $dangling nodes$\; 
 remove $dangling edges$\;
 \KwRet $WG$\;
 }
 \caption{Create a composition weighted graph from a queue.}
\label{graph_building}
\end{algorithm}

@@

\section{Results and Analysis}\label{results_analysis}
\subsection{Comparison Test with QoS Evaluation Model}\label{comparisonTest}
In this section, we analyse the composition solution generated by using our approach comparing with QoS-aware approach. First, we look at mean value of $Q_{mt}$, $Q_{s}$ and $Q_{QoS}$ at optimum at the 100th generation for two approaches, shown in Table \ref{decisionTable}. The QoS-aware approach record $Q_{mt}$, $Q_{s}$ and utilise fitness function $Fitness = w_1 \hat{A} + w_2 \hat{R} + w_3(1 - \hat{T}) + w_4(1 - \hat{C})$ when $\sum_{i=1}^{4} w_i = 1$, but $MT$ and $S$ are recorded, and $Q_(QoS)$ is normalised from 0 to 0.5 to make it comparable to $Q_{QoS}$ in our approach. We observe an interesting pattern from Table \ref{decisionTable} using statistic analysis: mean value of $Q_{mt}$ and $Q_{s}$ at optimum by our approach is consistently higher than those by QoS-aware approach.  Meanwhile, $Q_{QoS}$ generated by QoS-aware approach can obtain a slightly higher value than that of our approach. In conclusion, we can perceive that our evaluation could find out better functional quality with a reasonable trade off in QoS.
\begin{table}[]
\footnotesize
\centering
\caption{Mean Quality for comprehensive quality-aware methods and QoS-aware approach}
\label{decisionTable}
\begin{tabular}{l|l|l|l}
\hline
\multicolumn{2}{c|}{WSC09}              & \shortstack{QoS-aware \\ Evaluation} & \shortstack{Comprehensive \\ Quality Evaluation} \\ \hline
\multirow{3}{*}{Task1}  &$Q_{mt}$   &0.189787 $\pm$ 0.039278686            &0.221862 $\pm$ 0.009582 $\uparrow$ \\ \cline{2-4} 
                        &$Q_{s}$    &0.884962 $\pm$ 0.014140               &0.894082 $\pm$ 0.009206 $\uparrow$ \\ \cline{2-4}
                        &$Q_{QoS}$  &0.278730 $\pm$ 0.007786               &0.280222 $\pm$ 0.008212                          \\ \hline
\multirow{3}{*}{Task2}  &$Q_{mt}$   &0.001795 $\pm$ 0.000719               &0.001977 $\pm$ 0.001566 $\uparrow$ \\ \cline{2-4} 
                        &$Q_{s}$    &0.906971 $\pm$ 0.005855               &0.923970 $\pm$ 0.006898 $\uparrow$   \\ \cline{2-4}
                        &$Q_{QoS}$  &0.239979 $\pm$ 0.000578 $\uparrow$    &0.238596 $\pm$ 0.001264 \\ \hline
\multirow{3}{*}{Task3}  &$Q_{mt}$   &0.158526 $\pm$ 0.014028               &0.245830 $\pm$ 0.007761 $\uparrow$   \\ \cline{2-4} 
                        &$Q_{s}$    &0.949109 $\pm$ 0.002331               &0.972765 $\pm$ 0.002980 $\uparrow$           \\ \cline{2-4}
                        &$Q_{QoS}$  &0.247002 $\pm$ 0.000661 $\uparrow$    &0.245631 $\pm$ 0.000431            \\ \hline
\multirow{3}{*}{Task4}  &$Q_{mt}$   &0.000000 $\pm$ 0.000000               &0.000004 $\pm$ 0.000002 $\uparrow$ \\ \cline{2-4} 
                        &$Q_{s}$    &0.879514 $\pm$ 0.007456               &0.920733 $\pm$ 0.000001 $\uparrow$ \\ \cline{2-4}
                        &$Q_{QoS}$  &0.242297 $\pm$ 0.000507 $\uparrow$    &0.236677 $\pm$ 0.002211  \\ \hline
\multirow{3}{*}{Task5}  &$Q_{mt}$   &  0.000042 $\pm$ 0.000030             &0.000078 $\pm$ 0.000020 $\uparrow$           \\ \cline{2-4} 
                        &$Q_{s}$    &  0.915933 $\pm$ 0.012888             &0.927678 $\pm$ 0.002578 $\uparrow$                 \\ \cline{2-4}
                        &$Q_{QoS}$  & 0.238189 $\pm$ 0.000240 $\uparrow$      &  0.237485 $\pm$  0.000328  \\ \hline                                                   
\end{tabular}
\end{table}

>>>>

We defined our semantic web service composition solution as a weighted graph $WG = (V, E)$, where $V$ is a set of services as vertex: $V=\{S1, S2...Sn \}$ and $E$ is a set of edges $E = \{e_{1}, e_{2},... e_{n} \}$. Each edge is mapped to a semantic web service link, so each edge $e$ is associated with $sm_{L}$ as weight value as a pair value of quality values $mt_{L}$ and $s_{L}$ on the edge level, Here we provide an example of semantic web service composition, which is described in Fig. \ref{wscs}. The data of composition web service flows from the start to the end, where five web services involved and linked with each other using edge connections. Besides that, pair values that $mt_{L}$ and $s_{L}$ are calculated for all edges $e_{1}, e_{2},... e_{n}$ assigned as weights.

\begin{figure}[h]
\centering
\fbox{\includegraphics[scale=.28]{weigtedGraphExample.pdf}}
 \caption{Composition weighted graph solution}
 \label{wscs}
\end{figure}

@@

PSO has shown its efficiency in solving combinatorial optimisation problems \cite{fukuyama2008fundamentals}. Therefore, we will employ a PSO-based approach, which is considered to be simple and efficient without penalising or repairing that often required by GP \cite{da2016particle}. Fig. \ref{overview} shows the overview of our approach with five steps. Step 1: The composition process is triggered by a composition goal defined in Subsection \ref{problemDes}, which describes customers requirements both semantic matchmaking quality and QoS. Step 2: This composition goal are used to discover all relevant web services, which lead to a shrunken service repository that is subsequently used by PSO as a searching space. Step 3: A weighted graph representation is randomly built up from the an initial service queue that mapped to the particle's location, interleaving with semantic matchmaking process utilising ontology-based index cache. In the weighted graph, graph edges are assigned with semantic matchmaking quality as weights. Step 4: The fitness value of the weighted graph is evaluated to update the position of particle under PSO algorithm in Sect. \ref{pso_algorithm}, where the position is mapped to the index of service queue. later on, the updated service queue is used to decode a new weighted graph as the composition solution. Step 4. Lastly, the best position found in the searching space is selected and decode into the final optimised solution. This PSO-based approach is similar to \cite{da2016particle}, but we employ weighted graphs as a different solution presentation with edges transferred from semantic web service links.

\subsection{Ontology-based Index Cached Optimisation}\label{indexCache}
In PSO, particles represent weighted graphs with edges and vertices associated with quality of semantic web service matchmaking and QoS respectively, the bottlenecks of generating weighted graph lie in building edges and nodes, which are related to the cost of semantic quality calculation and the size of service repository respectively. To effectively construct weighted graphs, we pre-calculate semantic matchmaking quality. The key idea of the index is to create a mapping from the output-related concepts to potentially matched input-related concepts while considering different levels of match types. Meanwhile, this index size could also be reduced by only considering the concepts related in the shrunken service repository, which means the index cache is filtered by those task-relevant web services. This optimised cache also contributes to less and constant time for weighted graphs building through the whole evolutionary process.

@@

\begin{algorithm}
 \setlength\hsize{0.9\linewidth}
 \SetKwInOut{Input}{Input}\SetKwInOut{Output}{Output}
 \SetKwFunction{generateWeightedGraph}{generateWeightedGraph}
 \SetKwProg{Procedure}{Procedure}{}{}
 \LinesNumbered
 \SetNlSty{}{}{:}
 \Input{relevant Web services $rws$}
 \Output{Optimised service queue $queue$,Optimised Weighted Graph $OPTWG$}
 Map each relevant service to an index in the particle's position vector\;
 Randomly initialise each particle in the swarm\;
  \While {max. iterations not met}{
     \ForEach{particles in the swarm}{
     $queue \leftarrow$ map particle's position vector to service\;
     $WG \leftarrow$ $generateWeightedGraph$()\;
	 Calculate the $WG$ fitness value\;

      \eIf{fitness value better than \emph{pBest}}{    
        Assign current fitness as new \emph{pBest}\;
       }{
        Keep previous \emph{pBest}\;
       }	
     }
    Assign best particle's \emph{pBest} value to \emph{gBest}, if better than \emph{gBest}\;
 	Calculate the velocity of each particle\;
  	Update the position of each particle\;
  }
 \KwRet $OPTWG$\;
 \caption{Steps of the PSO-based Web service composition technique.}
\label{novelSteps}
\end{algorithm}

\begin{algorithm}
 \setlength\hsize{0.9\linewidth}
 \SetKwInOut{Input}{Input}\SetKwInOut{Output}{Output}
 \SetKwFunction{generateWeightedGraph}{generateWeightedGraph}
 \SetKwProg{Procedure}{Procedure}{}{}
 \LinesNumbered
 \SetNlSty{}{}{:}
  \Procedure{\generateWeightedGraph{}}{
 \Input{Task inputs $I$, task outputs $O$, Optimised service queue $queue$, IndexCache $IndexCache$}
 \Output{Weighted Graph $WG$}
 $WG \leftarrow  null$\;
 $WG \leftarrow$ new $endNode$(), new $startNode$()\;
 $OutputSet \leftarrow$ \{$I$\}\;
  \While {all $O$ $\notin$ $OutputSet$  and $queue !=null$}{
     \ForEach{$ws$ in $queue$}{
      \uIf{$ws.inputs$ $\in$ $OutputSet$ }{    
    		 \ForEach{$I$ in $ws.inputs$}{
    		$mt_{P}, s_{P} \leftarrow$ query $IndexCache$\;
     		$mt_{L} \leftarrow$ aggregation($mt_{P}$)\;
 			$s_{L} \leftarrow$ aggregation($s_{P}$)\;
              }
        $sm_{L} \leftarrow$ aggregation($mt_{L}$,$s_{L}$)\;
        $WG.edge \leftarrow$  $sm_{L} $\;
        $WG \leftarrow$ new $wsNode$()\;
         $OutputSet$ add \{$ws.outputs$\}\;    
       $queue$.remove $ws$\;
       }	
     }
  }
 remove $dangling nodes$\; 
 remove $dangling edges$\;
 \KwRet $WG$\;
 }
 \caption{Create a composition weighted graph from a queue.}
\label{graph_building}
\end{algorithm}

@@

\section{Results and Analysis}\label{results_analysis}
\subsection{Comparison Test with QoS Evaluation Model}\label{comparisonTest}
In this section, we analyse the composition solution generated by using our approach comparing with QoS-aware approach. First, we look at mean value of $Q_{mt}$, $Q_{s}$ and $Q_{QoS}$ at optimum at the 100th generation for two approaches, shown in Table \ref{decisionTable}. The QoS-aware approach record $Q_{mt}$, $Q_{s}$ and utilise fitness function $Fitness = w_1 \hat{A} + w_2 \hat{R} + w_3(1 - \hat{T}) + w_4(1 - \hat{C})$ when $\sum_{i=1}^{4} w_i = 1$, but $MT$ and $S$ are recorded, and $Q_(QoS)$ is normalised from 0 to 0.5 to make it comparable to $Q_{QoS}$ in our approach. We observe an interesting pattern from Table \ref{decisionTable} using statistic analysis: mean value of $Q_{mt}$ and $Q_{s}$ at optimum by our approach is consistently higher than those by QoS-aware approach.  Meanwhile, $Q_{QoS}$ generated by QoS-aware approach can obtain a slightly higher value than that of our approach. In conclusion, we can perceive that our evaluation could find out better functional quality with a reasonable trade off in QoS.
\begin{table}[]
\footnotesize
\centering
\caption{Mean Quality for comprehensive quality-aware methods and QoS-aware approach}
\label{decisionTable}
\begin{tabular}{l|l|l|l}
\hline
\multicolumn{2}{c|}{WSC09}              & \shortstack{QoS-aware \\ Evaluation} & \shortstack{Comprehensive \\ Quality Evaluation} \\ \hline
\multirow{3}{*}{Task1}  &$MT$   &0.189787 $\pm$ 0.039278686            &0.221862 $\pm$ 0.009582 $\uparrow$ \\ \cline{2-4} 
                        &$S$    &0.884962 $\pm$ 0.014140               &0.894082 $\pm$ 0.009206 $\uparrow$ \\ \cline{2-4}
                        &$QoS$  &0.278730 $\pm$ 0.007786               &0.280222 $\pm$ 0.008212                          \\ \hline
\multirow{3}{*}{Task2}  &$MT$   &0.001795 $\pm$ 0.000719               &0.001977 $\pm$ 0.001566 $\uparrow$ \\ \cline{2-4} 
                        &$S$    &0.906971 $\pm$ 0.005855               &0.923970 $\pm$ 0.006898 $\uparrow$   \\ \cline{2-4}
                        &$QoS$  &0.239979 $\pm$ 0.000578 $\uparrow$    &0.238596 $\pm$ 0.001264 \\ \hline
\multirow{3}{*}{Task3}  &$MT$   &0.158526 $\pm$ 0.014028               &0.245830 $\pm$ 0.007761 $\uparrow$   \\ \cline{2-4} 
                        &$S$    &0.949109 $\pm$ 0.002331               &0.972765 $\pm$ 0.002980 $\uparrow$           \\ \cline{2-4}
                        &$QoS$  &0.247002 $\pm$ 0.000661 $\uparrow$    &0.245631 $\pm$ 0.000431            \\ \hline
\multirow{3}{*}{Task4}  &$MT$   &0.000000 $\pm$ 0.000000               &0.000004 $\pm$ 0.000002 $\uparrow$ \\ \cline{2-4} 
                        &$S$    &0.879514 $\pm$ 0.007456               &0.920733 $\pm$ 0.000001 $\uparrow$ \\ \cline{2-4}
                        &$QoS$  &0.242297 $\pm$ 0.000507 $\uparrow$    &0.236677 $\pm$ 0.002211  \\ \hline
\multirow{3}{*}{Task5}  &$MT$   &  0.000042 $\pm$ 0.000030             &0.000078 $\pm$ 0.000020 $\uparrow$           \\ \cline{2-4} 
                        &$S$    &  0.915933 $\pm$ 0.012888             &0.927678 $\pm$ 0.002578 $\uparrow$                 \\ \cline{2-4}
                        &$QoS$  & 0.238189 $\pm$ 0.000240 $\uparrow$      &  0.237485 $\pm$  0.000328  \\ \hline                                                   
\end{tabular}
\end{table}
